---
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

# dataitd24: FAIRqual Workshop Data from ITD24

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.17473502.svg)](https://zenodo.org/doi/10.5281/zenodo.17473502)

<!-- badges: end -->

## About the Dataset

This repository contains the dataset from the FAIRqual workshop
conducted at the International Transdisciplinary Conference (ITD24) in
Utrecht, Netherlands (November 4-8, 2024). The workshop aimed to gather
experiences and perspectives on sharing qualitative data in
transdisciplinary research contexts.

### The FAIRqual Project

FAIRqual is a collaborative project between data stewards and qualitative researchers that aims to
develop both technological implementations and conceptual procedures
for sharing qualitative data based on the FAIR principles (Findable,
Accessible, Interoperable, Reusable) in transdisciplinary research.
This project is supported by the [Open Research Data Program of the ETH Board](https://ethrat.ch/en/eth-domain/open-research-data/).

Project website: <https://fairqual.org/about>

## Workshop Overview

The workshop brought together 24 researchers interested in
transdisciplinary research to:

- Share past experiences with open qualitative data
- Brainstorm challenges and opportunities for data sharing
- Imagine future scenarios for FAIR qualitative data practices
- Identify key concerns about sensitive data handling

Participants worked in four groups using flipcharts and post-it notes
to document their discussions. Key themes that emerged included:

- Handling of sensitive data and maintaining confidentiality
- Risks of data misuse and decontextualization
- Concerns about AI applications and data extraction
- Need for appropriate data repositories for qualitative data
- Cost-benefit considerations of preparing data for sharing

```{r, include=FALSE}
## Load packages needed for download tables
library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(tibble)
```

## Download

The workshop summary is available both as an online article and as a
downloadable DOCX file. The datasets (codebook and flipcharts) can be
downloaded as CSV or XLSX files from the tables below. For direct
programmatic access to the data in R, install the data package following
the guidance in the Installation section.

**Workshop Summary Document:**

```{r, echo=FALSE, message=FALSE, warning=FALSE}

extdata_path <- "https://github.com/fairqual/dataitd24/raw/main/inst/extdata/"
vignette_path <- "https://fairqual.github.io/dataitd24/articles/"

# Summary document table
summary_table <- tibble(
  Document = "summary_fairqual_workshop_itd24",
  Article = paste0("[Read Article](", vignette_path,
                   "summary-fairqual-workshop-itd24.html)"),
  Download = paste0("[Download DOCX](", extdata_path,
                    "summary_fairqual_workshop_itd24.docx)")
)

summary_table |>
  knitr::kable()

```

**Datasets:**

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Data files table
pictures_path <- "https://github.com/fairqual/dataitd24/tree/main/inst/extdata/pictures_flipcharts_fairqual_workshop_itd24"

data_files <- read_csv("data-raw/dictionary.csv") |>
  distinct(file_name) |>
  dplyr::mutate(file_name = str_remove(file_name, ".rda")) |>
  dplyr::rename(Dataset = file_name) |>
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, Dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, Dataset, ".xlsx)"),
    Pictures = case_when(
      Dataset == "flipcharts1" ~ paste0("[View Pictures](", pictures_path, ")"),
      Dataset == "flipcharts2" ~ paste0("[View Pictures](", pictures_path, ")"),
      TRUE ~ NA_character_
    )
  )

data_files |>
  knitr::kable()

```

## Installation

You can install the development version of dataitd24 from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("fairqual/dataitd24")
```

```{r}
## Run the following code in console if you don't have the packages
## install.packages(c("dplyr", "knitr", "readr", "stringr", "gt", "kableExtra", "tibble"))
library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(gt)
library(kableExtra)
library(tibble)
```



## Data

The package provides access to three main datasets from the FAIRqual
workshop:

```{r}
library(dataitd24)
```

### summary_fairqual_workshop_itd24

This document summarizes the workshop. It is available both as a
downloadable DOCX file and as an article on this website. The file
`summary_fairqual_workshop_itd24` describes the codes / subcategories /
categories that were identified in a qualitative content analysis
performed on all workshop materials (post-it notes from flipcharts; see
`flipcharts1` and `flipcharts2`). The aim of this document is to provide
context for the flipcharts and post-it notes created during the
workshop, as well as summarizing the discussions that took place. For a
tabular overview of the codes / subcategories / categories see the
dataset `codebook_qualitative`.

### codebook_qualitative

The dataset `codebook_qualitative` contains the codes / subcategories /
categories that were identified in a qualitative content analysis
performed on all workshop materials (post-it notes from flipcharts; see
`flipcharts1` and `flipcharts2`). The analysis categorized discussions
about sharing qualitative data in transdisciplinary research into
hierarchical themes. It has `r nrow(codebook_qualitative)` observations
and `r ncol(codebook_qualitative)` variables.


```{r}
codebook_qualitative |> 
  head(3) |> 
  gt::gt() |>
  gt::as_raw_html()
```

For an overview of the variable names, see the following table.

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "codebook_qualitative.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```

### flipcharts1

The dataset `flipcharts1` contains digitized post-it notes from the
first part of the workshop where participants shared lived and imagined
experiences with qualitative data sharing. It has
`r nrow(flipcharts1)` observations and `r ncol(flipcharts1)` variables.

```{r}
flipcharts1 |> 
  head(3) |> 
  gt::gt() |>
  gt::as_raw_html()
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "flipcharts1.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```

### flipcharts2

The dataset `flipcharts2` contains data from the second part of the
workshop where groups were combined to identify key challenges in
sharing qualitative data and propose approaches to address them. It has
`r nrow(flipcharts2)` observations and `r ncol(flipcharts2)` variables.

```{r}
flipcharts2 |> 
  head(3) |> 
  gt::gt() |>
  gt::as_raw_html()
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "flipcharts2.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```

## Ethics and Consent

Workshop participants were informed about data collection and provided
consent for sharing. Sensitive content could be marked with special
stickers, and participants had the option to opt out of photography
using conference-provided orange wristbands.

## License

Data are available as
[CC-BY](https://github.com/fairqual/dataitd24/blob/main/LICENSE.md).

## Citation

Please cite this package using:

```{r}
citation("dataitd24")
```

## Project Team

- [Dr. Franziska Mohr](https://tdlab.usys.ethz.ch/team/person-detail.franziska-mohr.html)
  ([USYS Transdisciplinarity Lab](https://tdlab.usys.ethz.ch/), ETH
  Zurich)
- [Lars Schöbitz](https://www.lse.de)
  ([Global Health Engineering](https://ghe.ethz.ch/), ETH Zurich)
- [Dr. Mollie Chapman](https://usys.ethz.ch/en/people/profile.MTY5Mjk4.TGlzdC82MzcsMzIwMTk3MjIy.html)
  ([USYS Transdisciplinarity Lab](https://tdlab.usys.ethz.ch/), ETH
  Zurich)
- [PD Dr. Bianca Vienni-Baptista](https://tdlab.usys.ethz.ch/team/person-detail.bianca-vienni.html)
  ([USYS Transdisciplinarity Lab](https://tdlab.usys.ethz.ch/), ETH
  Zurich)
- [Prof. Dr. Elizabeth Tilley](https://ghe.ethz.ch/about/people/person-detail.MTc2MjI5.TGlzdC80MTI2LC0xNDYwMDMwNTU3.html)
  ([Global Health Engineering](https://ghe.ethz.ch/), ETH Zurich)

## Acknowledgments

We thank all workshop participants for their valuable contributions
and insights. This work is supported by the Open Research Data Program
of the ETH Board.
